---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(rando)
library(tidyverse)
```

# rando

<!-- badges: start -->
<!-- badges: end -->

The goal of rando is to provide easier generating of random numbers 
in a manner that is context aware, and reproducible.

## Installation

You can install the development version of rando from 
[Github](https://github.com/MyKo101/rando) with:

``` 
install.packages("remotes")
install_github("MyKo101/rando")
```

Once installed, to load `rando`, use

```
library(rando)
```

## Example

With rando, generating random numbers becomes incredibly easy, as we 
do not need to define how many random numbers we need. `rando` will 
figure out how many you need based on where the number generator is
being used.

This works for `tibble()` declarations

```{r in-tibble}
df <- tibble(id = 1:10,
       x = r_norm())
df
```

and inside of `dplyr` verbs

```{r in-mutate}
mutate(df, y = r_unif())
```


Parameters can also be used to define the number of values to return.
If parameters are longer than 1, `rando` will try to return the same
number of random values, unless there is a clash between two of the
parameters

```{r argument-lengths, error=TRUE}
r_norm(mean = 1:10)
r_norm(mean=1:10,sd=1:2)
```
If you want to manually define the number of random numbers to be 
generated, there are two ways to do it. The old fashioned way: 
providing the `n` argument (this must be named)

```{r explicit-n}
r_unif(n=20)
```

Or, if we are generating many random numbers, we can set a default
`n` value to be used globally

```{r implicit-n}
set_n(15)
r_norm(mean=3)
r_binom(size=3)
```

## Safer and replicable
The `rando` functions also check if parameters being supplied are viable
and throws an informative error if not. This is different to the 
default `stats`  random number generating functions, which may return
a lot of `NaN` values with only a vague warning.

```{r sd-error, error=TRUE}
rnorm(n=10,sd=-1)
r_norm(sd=-1)
```

All `rando` functions can also take a `.seed` argument which does one
of two things:

* If a numeric is supplied, then `rando` will set this as the random 
seed before generating the values
* If a TRUE is supplied, then `rando` will randomly generate a 
numeric value to be used.

If `.seed` is not `NULL` (the default), then this `seed` value 
(supplied or generated) will be attached to the output, 
and can be extracted with `pull_seed()`

This allows for greater replicability in results.
```{r replicable}
r_norm(.seed = 42)
r_norm(.seed = 42)

x <- r_norm(.seed=TRUE)
x

r_norm(.seed=pull_seed(x))

```
## Blueprints

In order to make simulations easier, `rando` provides the `blueprint()` 
function. This function creates a plan for a simulated
dataset using `rando` functions.

```{r blueprint1}
make_tbl <- blueprint(
  x = r_norm(),
  y = r_norm()
)

make_tbl(n=2)

make_tbl(n=5)
```

These blueprints can accept additional arguments and will be generated
based on these arguments

```{r}
make_tbl2 <- blueprint(
  x = r_norm(mean=x_mu),
  y = r_unif(min=y_min,max=y_max)
)

set_n(10000)
make_tbl2(x_mu = 10, y_min = -10, y_max=-5) %>%
  summarise(n = n(), mean_x = mean(x), min_y = min(y), max_y = max(y))
```

This then allows for quick generation of simulation data using `pmap()`
and analysis using `map()`

```{r}
make_sim <- blueprint(
  x = r_norm(mean = x_mu),
  y = r_norm(mean = 2*x+10, sd = 2)
)

tibble(x_mu = r_unif(n = 5, -10, 10)) %>%
  pmap(make_sim, n = 100) %>%
  map(lm, formula = y ~ x) %>%
  map_dfr(broom::tidy)

```

## Distribution Functions

The majority of random number generating functions from the `stats`
package have been translated into `rando` functions. Be sure to look
into the documentation for the `rando` functions you use, as some have 
re-parametrised. Functions names for transitioning from `stats` to 
`rando` generally follow the same naming convention, that is `r*()` 
becomes `r_*()`, e.g. `r_norm()` replaces `rnorm()`. The only 
exceptions are `r_tdist()` and `r_fdist()` to take over the roles of
`rt()` and `rf()`, respectively. `rando` also includes several new distributions such
as `r_bern()` and `r_letters()`.

## Arbitrary Distributions

The `r_cdf()` function is a dynamic random number generator.
It can take any cdf as an argument and produce random numbers with the
associated distribution.

```{r r_cdf}

my_fun <- function(x,beta=1){
  if_else(x < 0, 0, 1-exp(-beta*x))
}

set_n(1000)
x_data <- r_cdf(my_fun)

hist(x_data,breaks=seq(0,10,0.1))

```
Any additional arguments used by the function, can be passed to
`r_cdf()`, and will be used in determining the number of values to
generate (just as in the other distribution functions above)

```{r}
r_cdf(my_fun,beta=1:10)
```

Finally, `purrr`-style functions can be used for `r_cdf()` to allow 
for even briefer function definitions. These have been extended to 
allow for the use of additional named arguments to be passed to these
`<lambda>` functions. Either `.x` or `.t` can be used for the random
variable.

```{r}
set_n(20)
r_cdf(~1-exp(-.x),min=0)

r_cdf(~1-exp(-beta*.x),beta=1:10,min=0,n=10)
```















